---
title: ''
author: ''
params:
  day: 2
  ptitle: yes
  pbackground: yes
  dtype: none
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
```

# Welcome to class! {data-background=#e8c35d}

## Reading Discussion

<br>

[The Data Science Process](https://medium.springboard.com/the-data-science-process-the-complete-laymans-guide-to-what-a-data-scientist-actually-does-ca3e166b7c67)

## RStudio Connect Problems

# Collaborating in GitHub {data-background=#e8c35d}

## How to write a GitHub issue

# Semester Project  {data-background=#e8c35d}

## Wednesday

<br>

Remember Wednesday is a work day. I will be on zoom (but not in the classroom) to answer questions during class.

## Presentation Outline

<br>

Friday we will **all meet on zoom**. In your breakout room, take turns presenting projects.

- Share your screen.
- You get 5 minutes to present.
  - Introduction/project goals
  - Where did you find your data?
  - Problems with data.
  - Wrangling steps. Did you learn anything new?
  - 
- Then 1 or 2 minutes to answer questions from the group.


## Next Week's Schedule

<br>

>- **Monday:**
>   - How to write issues in GitHub (part of the semester project)
>   - Presentations tips
>   - What else do you want to talk about??
>- **Wednesday:**
>   - Work day (no offical class)
>- **Friday:**
>   - Semester project presentations!



<!--------------------
## View Data at the Atomic Level

"None of us would admit now that we didn't realize...there...had to be spread in the data; we...just didn't realize how big it could be."

> - What key observations came by looking at the data
>   - small sample sizes
>   - values above 100%
>   - increasing # of patients at 100%
> - The Mean is Not the Distribution
-------------------------------------------->


<!------------------------
# Interactive Graphs {data-background=#e8c35d}

## Example: `library(DT)`

Make a new Rmd file that just has the following in it (including the default yaml). 

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval = FALSE}
library(DT)
datatable(iris)
```

Then, try knitting and pushing the files to GitHub

## Setting up R-Studio

> - Visit https://shiny.byui.edu/connect and log-in
> - Click on the publish button and choose "Manage Accounts"
> - Click the "connect" button and ensure the "Enable publishing to RStudio Connect" is checked
> - Use shiny.byui.edu as the server name.
> - Then knitting the .rmd file they click on the publish button

## Example: `library(DT)`

Look at the [DT: An R interface to the DataTables library](http://rstudio.github.io/DT/) website and find one or two modifications that you could make to your table output to move beyond the default.
-------------------------------->


<!-- Putting information in your file [My interactive](https://shiny.byui.edu/connect/#/apps/3/access) -->









# Your Future {data-background=#e8c35d}

## What are your goals?

- Where do you want to be in a week? A month? A year? 5 years? 10 years?
- How do you get there?


## Hustle culture

[Why are young people pretending to love work?](https://www.nytimes.com/2019/01/26/business/against-hustle-culture-rise-and-grind-tgim.html)

## [Workism Is Making Americans Miserable](https://www.theatlantic.com/ideas/archive/2019/02/religion-workism-making-americans-miserable/583441/?fbclid=IwAR1CiPDQJ8WpE7IyxwXBB_-cyy2gLo3Ecvhjbl3ude9_JRU5XcHTlcVtGKI)

<section style="text-align: left;">

- In the past century, the American conception of work has shifted from jobs to careers to callings-from necessity to status to meaning. In an agrarian or early-manufacturing economy, where tens of millions of people perform similar routinized tasks, there are no delusions about the higher purpose of, say, planting corn or screwing bolts: It's just a job.
- The rise of the professional class and corporate bureaucracies in the early 20th century created the modern journey of a career, a narrative arc bending toward a set of precious initials: VP, SVP, CEO. The upshot is that for today's workists, anything short of finding one's vocational soul mate means a wasted life.
- "We've created this idea that the meaning of life should be found in work," says Oren Cass, the author of the book The Once and Future Worker. "We tell young people that their work should be their passion. 'Don't give up until you find a job that you love!' we say. 'You should be changing the world!' we tell them. That is the message in commencement addresses, in pop culture, and frankly, in media, including The Atlantic."
- **But our desks were never meant to be our altars.**
- The problem with this gospel-Your dream job is out there, so never stop hustling-is that it's a blueprint for spiritual and physical exhaustion. Long hours don't make anybody more productive or creative; they make people stressed, tired and bitter.

</section>


## It's a Wonderful Life

## It's a Wonderful Life

**George:** I couldn't face being cooped up for the rest of my life in a shabby little office...I'd go crazy. I want to do something big and something important.

**Dad:** You know, George, I fell that in a small way we *are* doing something important.

## Personal Revelation

[Advice from the prophet.](https://www.churchofjesuschrist.org/study/general-conference/2018/04/revelation-for-the-church-revelation-for-our-lives?lang=eng)