---
title: "Leaflet Class Activity"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: show
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
library(USAboundaries)
library(tidyverse)
library(leaflet)
library(sf)
```

## Practice Using Leaflet for Interactive Maps 

- We will use leaflet to create interactive maps of the cumulative confirmed cases of COVID-19.
- First, we will plot the cities with the most confirmed cases.

Note, rather than doing copying and pasting, you may learn more by actually typing things out. This will force your mind to actually read over and digest the syntax.

Read in the COVID data. Replace the stars with yesterday's date m/d/y. This will give us a column called "confirmed" that contains the cumulative confirmed cases up to that date. (To learn more about the data, visit this [website](https://github.com/CSSEGISandData/COVID-19).)

```{r load_data}
# Use this R-Chunk to import all your datasets!
covid <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv") %>% 
  rename(confirmed = `*/*/*`)
```

## Plotting the Top Cities

We will plot just the top n cities by replacing the __ in top_n(). You get to decide how many cities to plot. (I'd recommend somewhere between 20 and 100). The code in the next chunk selects just those cities.

Don't forget to fill in the other blank in the code below.

```{r tidy_data}
# Use this R-Chunk to clean & wrangle your data!
covid_top_cities<- covid %>% arrange(____(confirmed)) %>% top_n(__) %>% select(confirmed, everything())
```


Now create a plot with a marker placed in the location of each of the top cities.
Note the use of the ~ in order to reference a column name.

```{r}
leaflet(data=covid_top_cities) %>% addTiles() %>% 
  addMarkers(lng = ~Long_, lat = ~Lat)
```

Consider the next chunk of code where we set some defaults for the map. What is this setView() doing and what values should you put in the blanks?

```{r}
leaflet(data=covid_top_cities) %>% addTiles() %>% 
  setView(lng = ___, lat = ___, zoom = ___) %>% 

```

Let's add labels to the markers so we know the name of the city and the cumulative # of confirmed cases. In addition to blanks, this code has 3 errors in it. One will prevent the code from running, the others are just mistake. Can you find and correct them?

```{r plot_data}
# Use this R-Chunk to plot & visualize your data!
leaflet(data=covid_top_cities) %>% addTiles() %>% 
  setView(lng = ___, lat = ___, zoom = ___) %>%  
  addMarkers(lng = ~Long_, lat = ~Lat, label = paste(Admin2, "confirmed" , "="))

```

## Circles on the map

Instead of markers we would like to use circles on the map. Let's encode the number of cases into the size of the circles.

```{r}
leaflet(data=covid_top_cities) %>% addTiles() %>% 
  setView(lng = -99, lat = 40, zoom = 4) %>%  
  addMarkers(lng = ~Long_, lat = ~Lat, label = ~paste(Admin2, "confirmed" , "=", confirmed))
```

## Your Turn

### Group 1: Circle Markers

Your job is to turn the markers in the above example into circles. Try changing the size of the circles to match the numbers of COVID cases. Use this website for help: [leaflet markers](https://rstudio.github.io/leaflet/markers.html).


```{r}
leaflet(data=covid_top_cities) %>% addTiles() %>% 
  setView(lng = -99, lat = 40, zoom = 4) %>% 

  ____(lng = ~Long_, lat = ~Lat, 
        _____ = ______,
        label = __________________)
```


In addition to circle size, we will encode the cumulative confirmed case count into the color of the circle. This is not as simple and straight forward as one would like. If you reach this point, come back to the main room to visit this site together: https://rstudio.github.io/leaflet/colors.html

First the set-up

```{r}
  addCircles(lng = ~Long_, lat = ~Lat, 
        radius = ~sqrt(confirmed)*200,
        label = ~paste(Admin2, "confirmed" , "=", confirmed))
```

### Group 2: Labels and Popups

Your job is to explain the difference between leaflet "labels" and "popups". Try adding a popup that appears over the Salt Lake City temple and says "This is the place." Use this website for help: [popups and labels](https://rstudio.github.io/leaflet/popups.html).

```{r}
leaflet(data=covid_top_cities) %>% addTiles() %>% 
  setView(lng = -99, lat = 40, zoom = 4) %>%  
  addMarkers(lng = ~Long_, lat = ~Lat, label = ~paste(Admin2, "confirmed" , "=", confirmed)) %>% 
  addPopups(lng = -111, lat = 40, popup = "This is the place.")
```

### Group 3: Polygon Boundaries

Your job is to add state boundaries to the map. We can already see the state boundaries as part of the background, but we'd like to add them as a leaflet layer that can be manipulated. Use the `USAboundaries` package to get the appropriate SHP file. Then use this website to figure out what layer to add: [lines and shapes](https://rstudio.github.io/leaflet/shapes.html).

```{r}
states48 <- USAboundaries::us_states() %>%  
  filter(name != "Alaska" & name != "Hawaii" & name != "Puerto Rico")

leaflet(data=covid_top_cities) %>% addTiles() %>% 
  setView(lng = -99, lat = 40, zoom = 4) %>%  
  addMarkers(lng = ~Long_, lat = ~Lat, label = ~paste(Admin2, "confirmed" , "=", confirmed)) %>% 
  addPolygons(data = states48)
```


## Coloring your markers

In addition to circle size, we will encode the cumulative confirmed case count into the color of the circle. This is not as simple and straight forward as one would like. https://rstudio.github.io/leaflet/colors.html

First, the set-up:


```{r}
pal<-colorNumeric(palette = c("yellow", "red"), domain = log(covid_top_cities$confirmed))

pal #this is to see what pal contains/is
```

Take the time to answer the following questions about the above line of code:

- What is the domain argument providing? What is it used for?
- Why are we taking the log do you think (try the plot without logging it first)
- What is pal?

Now to actually create the plot. Can you find the 2 errors?

```{r}
leaflet(data=covid_top_cities) %>% addTiles() %>% 
  setView(lng = -99, lat = 40, zoom = 4) %>% 
  addCircles(lng = ~Long_, lat = ~Lat, 
             radius = ~confirmed*20,
             label = ~paste0(Admin2," County = ", confirmed , " confirmed cases"), 
             fillOpacity = .1,
             color = ~pal((confirmed)),  #color is the border color
             fillColor =  pal(log(confirmed))) #notice the domain above must match exactly what is fed to pal
```


## Big Data and Spatial Data

Pause to read the following article: Spatial Data Often Becomes Big Data
https://blogs.esri.com/esri/arcgis/2017/10/17/strategies-to-effectively-display-large-amounts-of-data-in-web-apps/

Identify the 4 stratagies for displaying map data. Consider recording them in your Readme.md

1. 
2. 
3. 
4. 

```{r}
             radius = ~sqrt(confirmed)*200,
             label = ~paste0(Admin2," County = ", confirmed , " confirmed cases"), 
             fillOpacity = .3,
             color = ~pal((confirmed)),
             fillColor =  ~pal(log(confirmed)))
```


## Clustered markers and layers

Let's try clustered markers of the observations. This approach will allow us to include all the cities, not just the top ones.

Quickly scan the "Markers" section of the leaflet guide https://rstudio.github.io/leaflet/markers.html
to find the code you will need to fill in the blanks in the next R chunk.
```{r}
leaflet() %>% addTiles() %>% 
  addMarkers(data = covid,
             lng = ~Long_,
             lat = ~Lat,
             label = ~paste(Admin2, confirmed, sep = "="),
             _______ = _______)
```

The clustering isn't my favorite because it doesn't follow geopolitical boundaries. At the very least, let's plot those boundaries. 

```{r}
#First get the sf files that contain the boundaries
counties <- USAboundaries::us_counties() %>%  
  filter(name != "Alaska" & name != "Hawaii" & name != "Puerto Rico")
state48 <- USAboundaries::us_counties() %>%  
  filter(name != "Alaska" & name != "Hawaii" & name != "Puerto Rico")

#Now use the data to create the plot
leaflet() %>% addTiles() %>% 
  ___________(data = counties, fill = FALSE, weight = 1) %>%
  ___________(data = states48, fill = FALSE, weight = 2, color = "black") %>% 
  addMarkers(data = covid,
             ________ = _____________,
             lng = ~Long_,
             lat = ~Lat,
             label = ~paste(Admin2, confirmed, sep = "="))

```



```{r}
             clusterOptions = markerClusterOptions())
```

## Your Turn


Time permitting, read through the [Leaflet guide](https://rstudio.github.io/leaflet/) and find 1-2 additional map features you would like to play with.